REDO Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution I: Methods and Test Case

### 2017-11-29 main.tex
up to 5.2.2

\subsection{Chemical Evolution} 
\label{sec:chemical evolution}
Observations of Local Group dwarf galaxies have shown that they retain very little of the total amount of metals they are expected to have produced, given their star formation histories (see Section~\ref{sec:observation}). In Figure~\ref{fig:species_fractions} we show the fraction of metals retained within the disk / expelled into the halo of our galaxy over time, along with a break down of the locations of each metal species by the end of the simulation time. As shown in the top panel, a majority of each metal tracer species is found outside the disk of the galaxy, in the galaxy halo. In total, \textbf{85\%} of the tracer metal species produced lie outside of the disk of our galaxy. Interestingly, the exact fraction is not uniform across metal species. Most notably, Nitrogen, is preferentially retained by the galaxy over the other species.

Roughly \textbf{35\%} of the N produced remains in the disk, as compared to \textbf{15\%} of O and \textbf{15\%} of Fe, for example. This difference is entirely feedback driven, but a coincidence of our particular galaxy conditions. At our galaxy's metallicity, the IMF weighted production of N rests entirely with the stellar winds of stars between 8 and 10 M$_{\odot}$; there is little to no N production in supernovae at this metallicity. The rest of the species either have very little stellar wind production, produced solely in supernovae, or have similar production levels between both supernovae and stellar winds (like O, for example). The typically lower injection energies of N into the galaxy than the SN injected elements means N is preferentially retained within the ISM, and not as effectively ejected through outflows. In truth, the stars that produce N do eventually explode as supernovae, sweeping up some fraction of the wind produced N into outflows. However, clearly some of this N is able to mix more effectively within the ISM than the other metal species. How abundance variations like this evolve over longer timescales and with varying feedback physics is the subject of ongoing work (Emerick et. al. in prep.).

\begin{figure*}
\includegraphics[width=0.9\linewidth]{species_bar.png}\\
\includegraphics[width=0.9\linewidth]{species_bar_disk.png}
\caption{The fraction of each species in the full simulation box contained in the halo, gas in the galaxy disk, and stars (top), and the fraction of species within the disk alone each in phase of the ISM (bottom).}
\label{fig:species_fractions}
\end{figure*}

The bottom panel illustrates where these metals are partitioned within the ISM of the galaxy at the end of the simulation time. The CNM contains roughly 50-60\% of each species, with the CNM and WNM together accounting for most of the metals in the galaxy. The WIM contains only $\sim$ 10\% of each species, with the HIM containing very little. This may be counterintuitive to the fact that most metals are initially injected into the galaxy in either the WIM or HIM phases. However, the disk metals represent only a small fraction of the total metals produced. Much of the metals are effectively removed from the disk into the halo (see Section~\ref{sec:outflows}). It is apparently that the remaining metals quickly mix or cool into denser gas phases.

@@

The observational sample of isolated, gaseous, low mass dwarf galaxies is limited compared to more massive galaxies, but has improved substantially with recent blind and targeted \ion{H}{1} surveys \citep[e.g.][]{Giovanelli2005, Geha2006, Geha2012, Walter2008, Cannon2011, Haynes2011, Hunter2012, Bradford2015, Tollerud2015, Sand2015}. However, the sample of isolated dwarf galaxies with $M_{*} < 10^{7}$~M$_{\odot}$ remains small. In Figure ~\ref{fig:KS} we show where our galaxy lies relative to the observed Kennicutt-Schmidt (KS) law (left) for low mass galaxies, and the extended Schmidt law (ESL) (right) which has been shown recently to be extend to low mass, low metallicity dwarf galaxies \citep{Roychowdhury2017}. We plot our galaxy every 1 Myr in these diagrams, showing the significant temporal variation in both spaces. Although these quantities are conceptually simple, due to how these values are measured observationally it is challenging to make a direct comparison to simulations, particularly in defining the geometric region to measure the surface densities. We have attempted to make a reasonable analog to how $\Sigma_{\rm sfr}$ and $\Sigma_{\rm gas}$ are measured observationally for low mass dwarfs following what is described in \citet{Roychowdhury2014}. We define $\Sigma_{\rm sfr} = \dot{M}_{*, \rm{10 Myr}} / A_{*, \rm{10 Myr}}$, where $\dot{M}_{*, \rm{10 Myr}}$ is the SFR measured over the preceding 10 Myr, and $A_{*, \rm{10 Myr}}$ is the disk defined out to the farthest star formed within the previous 10 Myr. Likewise, $\Sigma_{\rm gas} = M_{\rm gas, 10 Myr} / A_{*, \rm{10 Myr}}$, where M$_{\rm gas, 10 Myr}$ is the total gas mass within this defined disk. However, the total gas content cannot be determined observationally, particularly for low mass dwarf galaxies with a virtually unmeasurable H$_2$ content. Therefore, we present $\Sigma_{\rm gas}$ as measured directly in the simulations (green/purple points) and the observational equivalent, or $\Sigma_{\rm gas, obs} = 1.34 \times \Sigma_{\rm HI}$, where the factor 1.34 attempts to account for He, and where there is no correction for H$_2$ (plasma points).

@@

In both panels, we also include recent observational constraints from these relationships. We include the KS fits from \citet{Shi2011} with 1$\sigma$ errors (black), the fits from \citet{Roychowdhury2014} (blue), and the individual points from the recent SHIELD sample \citep{Teich2016}; we include only the fit and 1$\sigma$ errors from \citep{Shi2011} for the ESL (black), which have been shown to match well for dwarf galaxies at these surface densities \citep{Roychowdhury2017}. In both relationships our simulations fit well around the median observed fit, spanning the 1 $\sigma$ variations about the mean. This is a remarkable result. In constructing our galaxy model, we employed \textit{no} tuning of the underlying physics, adopting only canonical values for any available free parameters. It is non-trivial that our galaxy should oscillate about the median relationships in Fig.~\ref{fig:KS}, but signifies a proper accounting of the relevant physics governing gas and star forming properties in our galaxy. This result is consistent with recent galactic evolution simulations run at high resolution with a detailed accounting of stellar feedback physics (\textit{relevant citations})

>>>>

\subsection{Chemical Evolution} 
\label{sec:chemical evolution}
Observations of Local Group dwarf galaxies have shown that they retain very little of the total amount of metals they are expected to have produced, given their star formation histories (see Section~\ref{sec:observation}). In Figure~\ref{fig:species_fractions} we show the fraction of metals retained within the disk or expelled into the halo of our galaxy over time, along with a breakdown of the locations of each metal species by the end of the simulation. As shown in the top panel, a majority of each metal tracer species is found outside the disk of the galaxy, in the galaxy halo. In total, \textbf{85\%} of the tracer metal species produced lie outside of the disk of our galaxy. Interestingly, the exact fraction is not uniform across metal species.

Most notably, N is preferentially retained by the galaxy over the other species. 
Roughly \textbf{35\%} of N produced remains in the disk, as compared to \textbf{15\%} of O 
%mm and \textbf{15\%} of 
    or
Fe, for example. This difference is entirely feedback driven, but 
%mm a coincidence of our particular galaxy conditions. At our galaxy's metallicity, the IMF weighted production of N occurs entirely
    determined by the chosen metallicity of our galaxy. At this metallicity, N is 
    produced only 
in the stellar winds of stars between 8 and 10 M$_{\odot}$, with little to no production in supernovae. The rest of the species either have very little stellar wind production, are produced solely in supernovae, or have similar production levels between both supernovae and stellar winds (like O, for example). The typically lower injection energies of N into the galaxy compared to the SN 
%mm injected 
   produced
elements means N is preferentially retained within the ISM, and not as effectively ejected through outflows. In truth, the stars that produce N do eventually explode as supernovae, sweeping up some fraction of the wind produced N into outflows. However, clearly some of this N mixes more effectively within the ISM than the other metal species. How abundance variations like this evolve over longer timescales and with varying feedback physics is the subject of ongoing work.
%mm (Emerick et. al. in prep.).

\begin{figure*}
\includegraphics[width=0.9\linewidth]{species_bar.png}\\
\includegraphics[width=0.9\linewidth]{species_bar_disk.png}
\caption{The fraction of each species in the full simulation box contained in the halo, gas in the galaxy disk, and stars (top), and the fraction of species within the disk alone in each phase of the ISM (bottom).}
\label{fig:species_fractions}
\end{figure*}

The bottom panel illustrates where these metals are partitioned within the ISM of the galaxy at the end of the simulation time. The CNM contains roughly 50-60\% of each species, with the CNM and WNM together accounting for most of the metals in the galaxy. The WIM contains only $\sim$ 10\% of each species, with the HIM containing very little. This may be counterintuitive given
that most metals are initially injected into the galaxy in either the WIM or HIM phases. However, the disk metals represent only a small fraction of the total metals produced. Much of the metals are effectively removed from the disk into the halo (see Section~\ref{sec:outflows}). It is apparent that the remaining metals quickly mix or cool into denser gas phases.

@@

The observational sample of isolated, gaseous, low mass dwarf galaxies is limited compared to more massive galaxies, but has improved substantially with recent blind and targeted \ion{H}{1} surveys \citep[e.g.][]{Giovanelli2005, Geha2006, Geha2012, Walter2008, Cannon2011, Haynes2011, Hunter2012, Bradford2015, Tollerud2015, Sand2015}. However, the sample of isolated dwarf galaxies with $M_{*} < 10^{7}$~M$_{\odot}$ remains small. In Figure ~\ref{fig:KS} we show where our galaxy lies relative to the observed Kennicutt-Schmidt (KS) law (left) for low mass galaxies, and the extended Schmidt law (ESL) (right) which has been shown recently to extend to low mass, low metallicity dwarf galaxies \citep{Roychowdhury2017}. We plot our galaxy every 1 Myr in these diagrams, showing the significant temporal variation in both spaces. Although these quantities are conceptually simple, due to how these values are measured observationally it is challenging to make a direct comparison to simulations, particularly in defining the geometric region to measure the surface densities. We have attempted to make a reasonable analog to how $\Sigma_{\rm sfr}$ and $\Sigma_{\rm gas}$ are measured observationally for low mass dwarfs following the procedure described in \citet{Roychowdhury2014}. We define $\Sigma_{\rm sfr} = \dot{M}_{*, \rm{10 Myr}} / A_{*, \rm{10 Myr}}$, where $\dot{M}_{*, \rm{10 Myr}}$ is the SFR measured over the preceding 10 Myr, and $A_{*, \rm{10 Myr}}$ is the disk defined out to the farthest star formed within the previous 10 Myr. Likewise, $\Sigma_{\rm gas} = M_{\rm gas, 10 Myr} / A_{*, \rm{10 Myr}}$, where M$_{\rm gas, 10 Myr}$ is the total gas mass within this defined disk. However, the total gas content cannot be determined observationally, particularly for low mass dwarf galaxies with a virtually unmeasurable H$_2$ content. Therefore, we present $\Sigma_{\rm gas}$ as measured directly in the simulations (green/purple points) and the observational equivalent, or $\Sigma_{\rm gas, obs} = 1.34 \times \Sigma_{\rm HI}$, where the factor 1.34 attempts to account for He, and where there is no correction for H$_2$ (plasma points).

@@

In both panels, we also include recent observational constraints from these relationships. We include the KS fits from \citet{Shi2011} with 1$\sigma$ errors (black), the fits 
%mm
   and errors 
from \citet{Roychowdhury2014} (blue), and the individual points 
%mm
   (black squares) 
from the recent SHIELD sample \citep{Teich2016}; we include only the fit and 1$\sigma$ errors from \citep{Shi2011} for the ESL (black), which have been shown to match well for dwarf galaxies at these surface densities \citep{Roychowdhury2017}. In both relationships our simulations fit well around the median observed fit, spanning the 1 $\sigma$ variations about the mean.

This is a remarkable result. In constructing our galaxy model, we employed \textit{no} tuning of the underlying physics, adopting only canonical values for any available free parameters. It is non-trivial that our galaxy should oscillate about the median relationships in Fig.~\ref{fig:KS}, but signifies a proper accounting of the relevant physics governing gas and star forming properties in our galaxy. This result is consistent with recent galactic evolution simulations run at high resolution with a detailed accounting of stellar feedback physics (\textit{relevant citations})
%mm 
   but also with the demonstration by \citet{Li2005} that the KS law can be reproduced 
   by gravitation acting on an isothermal disk, which was argued to demonstrate that 
   this represents the basic physics at work.
